# R package for working with Hansen et al. 2013 data

```{r setup_knitr_options, results='hide', echo=FALSE, cache=FALSE}
require(knitr)
# Setup reST options:
options(width = 75)
opts_chunk$set(cache=TRUE, cache.path='knitr-cache/',
               fig.path='figures/', fig.width=3, fig.height=3, out.width='3in', 
               out.height='3in', dpi=300)
```

## Overview
This `gfcanalysis` R package facilitates simple analyses using the Hansen et 
al.  2013 Global Forest Change dataset. 

See the help files for the functions below for more information. For example, 
type `?download_tiles` in R to see the help file for the "download_tiles" 
function.

NOTE: the `gfcanalysis` package must be installed before this script will run.  
Run the "install_gfcanalysis.R" script to install/update the `gfcanalysis` 
package.

Load the `gfcanalysis` package

```{r}
library(gfcanalysis)
```

Load 'rgdal' package, which is used to read/write shapefiles and rasters

```{r}
library(rgdal)
```

Load 'spatial.tools' package, which is used to process rasters in parallel

```{r}
library(spatial.tools)
```

Start the parallel processing engine - comment out this line, and the 
sfQuickStop() line at the end of this script if you do NOT want `gfcanalysis` 
to use all of the available computing power on your machine (for example if you 
need to use the computer for email, etc. while it is processing GFC data)

```{r}
sfQuickInit()
```

Indicate where we want to save GFC tiles downloaded from Google. For any given 
AOI, the script will first check to see if these tiles are available locally 
(in the below folder) before downloading them from the server - so I recommend 
storing ALL of your GFC tiles in the same folder.

```{r}
output_folder <- 'C:/Users/azvoleff/Desktop/gfcanalysis_demo'
```

Set the threshold for forest/non-forest based on the treecover2000 layer in 
the GFC product

```{r}
forest_threshold <- 90
```

## Downloading data from Google server for a given AOI

Load a demo AOI from the P drive - notice that first we specify the folder 
the shapefile is in, and then the name of the shapefile without the '.shp'

```{r}
aoi <- readOGR('P:/Alex Zvoleff/gfcanalysis_demo', 'ZOI_NAK_2012')
```

Calculate the google server URLs for the tiles needed to cover the AOI

```{r}
tiles <- calc_gfc_tiles(aoi)
```

Check to see if these tiles are already present locally, and download them if 
they are not.

```{r}
download_tiles(tiles, output_folder, first_and_last=FALSE)
```

Extract the GFC data for this AOI from the downloaded GFC tiles, mosaicing 
multiple tiles as necessary (if needed to cover the AOI).

```{r}
gfc_data <- extract_gfc(aoi, output_folder)
```

Save the output data to a GeoTIFF (can also save in ENVI format, Erdas 
format, etc.)

```{r}
writeRaster(gfc_data, filename='test_gfc_extract.tif')
```

## Performing thresholding and calculating basic statistics

Calculate annual statistics on forest loss/gain

```{r}
gfc_stats <- gfc_stats(aoi, gfc_data, forest_threshold=forest_threshold)
```

Save these statistics to CSV files for use in Excel, etc.

```{r}
write.csv(gfc_stats$loss_table, file='test_gfc_extract_losstable.csv', row.names=FALSE)
write.csv(gfc_stats$gain_table, file='test_gfc_extract_gaintable.csv', row.names=FALSE)
```

Calculate and save a thresholded version of the GFC product

```{r}
gfc_thresholded <- threshold(gfc_data, forest_threshold=forest_threshold)
writeRaster(gfc_thresholded, filename='test_gfc_extract_thresholded.tif')
```

## Making simple visualizations

Calculate and save a thresholded annual layer stack from the GFC product 
(useful for simple visualizations, etc.)

```{r}
gfc_thresholded_annual <- annual_stack(gfc_data, forest_threshold=forest_threshold)
writeRaster(gfc_thresholded_annual, filename='test_gfc_extract_thresholded_annual.tif')
```

Save a simple visualization of the thresholded annual layer stack (this is 
just an example, and is using the data in WGS84. The data should be projected 
for this).

```{r}
animate_annual(aoi, gfc_thresholded_annual)
```

Stop the parallel processing engine

```{r}
sfQuickStop()
```
